---
title: "Basic tokenization"
output: html_document
---

```{r}
  library(dplyr)
  library(here)
  library(tidytext)
  library(knitr)
  library(tidyr)
```


```{r}
# If the file of saved data is not present
if (!file.exists(here("data/quakersaints.rda"))) {
     # download the data from Project Gutenberg
     quakersaints <-  gutenberg_download(c(19605))
     # and save it to that file name.
     save(quakersaints, file= here("data/quakersaints.rda"))
# Otherwise, 
} else {
     # load the data set from the file.
     load(here("data/quakersaints.rda"))
}
```

Tokenization means breaking something down 
into parts.  When working with text data, that usually means breaking 
it into words, although sometimes it means using other units like 
sets of two words or sentences.

Let's start by looking at the top or head of the quakersaints data. This 
by default shows the first 5 rows of the data frame.
As you an see once it is run, there are two variables, 
gutenberg_id and text. (Since this data set only includes one book, the 
value of gutenberg_id is the same for all of the rows.)

```{r }
head(quakersaints)
```

Now we want to take this data set of 15205 rows, where each 
row is a line from the book, and divide them up into a data 
set with one word per line. 
```{r}
quakersaints |> 
  unnest_tokens(output=word, input = text, token = "words") ->
   qs_tokens
head(qs_tokens)
tail(qs_tokens)
```

Once we have this list of words, we can do a lot of interesting
things. 

The `unnest_tokens()` function, by default,
does a number of other things to our data. 
First, it converts all of the letters to lower case. 
Second, it strips most punctuation.  That is why the "----" lines
at the beginning are no longer there.

There are some basic things we can do, such as find out how many
unique words there are.

```{r}
unique(qs_tokens$word) |> length()
```

There are just over 10,000 unque words represented in the 141,639 
tokens.

We can also find the most common words.

```{r}
qs_tokens |> count(word, sort = TRUE) |> head(20)
```
The word "i" is used 1329 times and "it" is uses 1329 times. 

You might notice that the most common words are almost all pretty 
short and are words we would expect to see as most common in almost
any piece of English writing.

We can also make visualizations.  A simple one is to make a bar chart 
of the most common words. 

```{r}
qs_tokens |> count(word, sort = TRUE) |>
  filter(n > 600) |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

## Stop words

Common words like "the" and "is" overwhelm other words. While for 
some puposes they are important, for others they are not useful.
To address the issue of common words generally we use the 
idea of *stop words*. These are words that we will remove from 
our tokenized data. 

The tidytext package contains a list of English stop words in the 
form of a dataframe `stop_words`. 
We want to remove all the wordsin the stop word list from our qs_tokens data.
We can do is using an function called `anti_join()`.

```{r}
 qs_tokens |> anti_join(stop_words) -> qs_tokens2
unique(qs_tokens2$word) |> length()
nrow(qs_tokens2)
head(qs_tokens2)
```

In our new data set we are down to 50,173 tokens representing 9478
unique words. So it's almost as many words as before, but we've
eliminated a lot of tokens because we have taken out the 
extremely common words. 

```{r}
qs_tokens2 |> count(word, sort = TRUE)
```
Now we see that the most common words are a lot more interesting.


```{r}
qs_tokens2 |> count(word, sort = TRUE)
```

This time instead of making a graph with all the words that 
appear over 600 times, let's look at those that appear 
at least 100 times.

```{r}
qs_tokens2 |> count(word, sort = TRUE) |>
  filter(n > 100) |>
  mutate(word = reorder(word, n)) |>
  
  ggplot(aes(x = word, y = n)) +
  coord_flip() +
  geom_col() +
  labs(y = NULL)
```

This is a much more interesting list. 

This is a good start but as you can see if you look at the head of
qs_tokens2 there are still a number of issues. For example all of the
text that is about the Project Gutenberg process is in there, and we don't 
want to analyze that. 

So even though we have some basic data now, we will still 
benefit from some more preparation and curation.