---
title: "Curating the data"
output: html_document
---

```{r}
library(gutenbergr)
library(tidytext)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(scales)
library(ggforce)
library(forcats)
library(here)
```

I don't want to redownload each time I run this file (because it takes
time and is an unnecessary load on Project Gutenberg) . I can save a working copy of the data and use that
when experimenting.  Doing it this way I can still run the whole
process by deleting or renaming the "data/quakersaints.rda" file. 

To load the saved dataset I use the `load()` function.

```{r }
if (!file.exists(here("data/quakersaints.rda"))) {
     quakersaints <-  gutenberg_download(c(19605))
     save(quakersaints, file="data/quakersaints.rda")
} else {
     load(here("data/quakersaints.rda"))
}
```

Previously, we dove straight into tokenizing. But this time let's 
first think about how we could solve some issues. 

Looking specifically at this book, I notice first that there are the 
standard Project Gutenberg texts at the beginning, followed by an 
preface which ends on line 182.

Then from 192 to 260 there is a table of contents with many blank lines. 
Also * in the title indicates "for younger children." 

I would like to separate each of the stories

Using this information, I can subset the data.

```{r}

table_of_contents <- quakersaints[190:260,] 
# Remove blank lines
table_of_contents <- table_of_contents |> filter(text != "")
table_of_contents$text
```

What do you notice about the table of contents? 
I see at least 3 issues.

1. The * at the beginning of some lines, which indicate whether
the story is appropriate for younger people.

2. The titles are followed by white spaces and then a page number.

3. There are roman numerals, both for the chapter numbering and 
the page numbering for the preface.

```{r}
table_of_contents <- table_of_contents |>
                     mutate(younger = substr(text, 1, 1) == "*")
```


```{r}
table_of_contents <- table_of_contents |>
                     mutate(title = "Title")
```

```{r}
# Remove numbers (could also use "[0-9]+")
table_of_contents <- table_of_contents |> 
  mutate(text =gsub("[[:digit:]]+", "", text))
# Deal with a special case
table_of_contents <- table_of_contents |>
  mutate(text = sub("_page_ vii", "", text, fixed = TRUE))
# Remove *
table_of_contents <- table_of_contents |> 
  mutate(text = sub("*", "", text, fixed=TRUE))
# Remove extra white space
table_of_contents <- table_of_contents |> 
  mutate(text =trimws(text))
table_of_contents$text
```

Notice that without the last step there is a lot of "white space" around the titles.  I
assume that these are " " values rather than tabs or other white 
space characters.  However if that turns out to be wrong I will use
one of the functions from stringr to get rid of them.

All of the above mutates could be combined into one `mutate()` to be more
efficient. They are separate to illustrate the step by step process.

Also 
 [1] "PREFACE"                                                
 [2] "A TALK ABOUT SAINTS" 
 [35] "COME-TO-GOOD" 
 [36] "HISTORICAL NOTES"  
 are different than the other sections so we may want to come back
 and handle them. They don't have chapter numbers which indicates
 that they are different kinnds of text.

I'm going to cut down the quakersaints data so that it just contains
text after the table of contents and list of illustrations. 

```{r}
quakersaints_trimmed <- quakersaints[290:nrow(quakersaints), ]
```

I'll also add a variable to represent the line number in the trimmed
data. 
As usual there are many ways to do this, I'll use the base method so
you see it , but `mutate(id = row_number())` will also work.

```{r}
quakersaints_trimmed <-  quakersaints_trimmed |> 
  mutate(row_number = row_number())
```

Let's find the chapter titles in the text.

In the table of contents data the text variable in 
that data is the same as the text variable in quakersaints_trimmed.
I can now find the row numbers (locations) of the start of each 
chapter by finding the row numbers of the matching text.
I use the `%in%` function to say if the text in a row matches 
one of the texts in the table of contents, keep it, otherwise
drop it.

```{r}
title_locations <- 
  quakersaints_trimmed |> 
  filter(text %in% table_of_contents$text)
title_locations
```


```{r}
table_of_contents |> left_join(quakersaints_trimmed, 
                               by = c("gutenberg_id", "text")) ->
  table_of_contents
```




What's the problem now?    
We can see that each title is repeated twice in the text.

Go back and look at the text.  You can see that each chapter
has some introductory items that are usually poems or quotations.
These types of material are called epigraphs. (epi = over or on, 
graph = writing)

This means that to understand our text, we would want to separate it 
into two groups (epigraph and story). I think we should also
mark text as either a title or a text and also mark the non-stories 
(the ones without Roman numerals) as different from the stories. 

So we need: Title versus text
            Epigraph versus body
            Front matter (Preface, A talk about saints) versus
                 Back matter (Come to Good, Historical notes) versus
                 Stories

There are many, many ways that we could do this. Figuring it out 
requires you to think through the steps logically. 
Assume that you will have to try many different logics to figure
this out.  Sometimes you can be clever and sometimes you can
be too clever, it takes practice to figure it out.  Each
text you work with 

We can use this information to say what kind of text each line is. 
The case_when() function is useful when we have complicated if then/else if/ 
else if .../else situations.

```{r}
title_locations <- 
  quakersaints_trimmed |> 
  mutate(text = trimws(text)) |>
  right_join(table_of_contents, by = c("text", "row_number",
                                       "gutenberg_id")) |>
  group_by(text) |> 
  mutate(toc_number =  row_number(),  
         text_type = case_when( 
                    text %in% 
                        c("PREFACE", "A TALK ABOUT SAINTS") ~ 
                                          "Front Matter Title",
                    text %in% c("HISTORICAL NOTES") ~ "Back Matter Title",
                    toc_number == 1 ~ "Epigraph Title", 
                    TRUE ~ "Story Title"),
         section_type = case_when(
           text_type == "Epigraph Title" ~ "Epigraph",
           text_type == "Story Title" ~ "Story",
           text_type == "Front Matter Title" ~ "Front Matter",
           text_type == "Back Matter Title"  ~ "Back Matter"
         )) 
title_locations
```



```{r}

```
So now I know where my stories and epigraphs start and the type of
text.   But 
I need to pull this back into the full data set and 
actually assign all of the lines to their story. 

As always there are a lot of ways to do that, but to keep it simple
we will use the `fill()` function.

Let's start by creating a variable indicating chapter. Then
we will fill that using the title.


```{r}

quakersaints_trimmed |> 
  left_join(title_locations, by = c("row_number", "gutenberg_id")) |>
  # We get text.y and text.x because text is in both data frames,
  # but not in the by vector.
  rename(title = text.y, text = text.x) |>
  mutate(text_type = case_when(is.na(text_type) ~ "Text",
                               TRUE ~ text_type)) |>
  fill(title) |> fill(section_type) |> fill(younger) -> quakersaints_trimmed

```
My analysis of the bigrams found that the most common ones were some names:

george fox	201			
robert fowler	36			
james nayler	31			
		
judge fell
john audland


meeting house	26	
steeple house 

Before tokenizing I am going to use stringr to insert underscores between the 
two words so that they are treated as one word.  


thou art
thou hast 
Add to stop words:

thou
thee
hast
art 

I am going to make those stop words

```{r}
stop_wordsq <- data.frame(word=c("thou", "thee", "thy", "art", "hast", "thine"))
```


In my case
I suspect the poetic language might be different so I have indicated for each
line whether it is part of an epigraph or story. 
I might also suspect that the language differs between the stories that are 
considered appropriate for younger children.

I may also want to eliminate the titles from my analyses.

So now I will tokenize.
```{r}
quakersaints_trimmed |> 
  filter(text_type == "Text") |>
  mutate(text = tolower(text)) |>
  # do this first because we will add underscores
  mutate(text = str_replace(text, "_", ""))|>
  # simple stemming to remove 's
  mutate(text = str_replace(text, "'s", ""))|>
  #mutate(text = str_extract(text, "[a-z]+")) |>
  mutate(text = str_replace_all(text,
                                c("george fox" = "george_fox",
                                  "g.bfox" = "george_fox",
                                  "robert fowler" = "robert_fowler",
                                  "james nayler" = "james_nayler",
                                  "judge fell" = "judge_fell",
                                  "john audland" = "john_audland",
                                  "meeting-house" = "meeting_house",
                                  "steeple-house" = "steeple_house",
                                  "valiant sixty" = "valiant_sixty",
                                  "edward burrough" = "edward_burrough",
                                  "e. burrough" = "edward_burrough"
                                  )))|>
  filter(!is.na(text))  |>
   unnest_tokens(word, text, to_lower = TRUE) ->
  quakersaints_long

```

This time I am going to remove the punctuation and numbers and _then_ remove
the stop words.

```{r}

quakersaints_long |>  
  #mutate(word = str_extract(word, "[a-z']+")) |>
  anti_join(stop_words) |>
  anti_join(stop_wordsq) |>
   filter(!is.na(word) )->
  quakersaints_long

```

Let's look at the top 20 words in the book as a whole. 

```{r}
quakersaints_long |> count(word, sort = TRUE) #|> head(50)
```

Looking these over, what should I drop?  How could I have done this 
earlier? 

```{r}
#quakersaints_long <- quakersaints_long |> filter (is.na(word) == FALSE)
#quakersaints_long |> count(word, sort = TRUE) |> head(20)
# Top 15 words for each type
quakersaints_long |> group_by(section_type) |> count(word, sort = TRUE) |>
  slice_head(n = 15)
# Calculate for each word, what proportion of words in a type it makes up.
quakersaints_long |> group_by(section_type) |> count(word, sort = TRUE) |>
  mutate(proportion = n / sum(n)) |> select(-n) |>
   pivot_wider(values_from = proportion, names_from = section_type) -> qs_props_section_type
```


```{r}
qs_props_section_type |> 
  mutate(anystory = !is.na(Story), anyepigraph = !is.na(Epigraph)) |>
  group_by(anystory, anyepigraph) |>
  summarize(n = n())
```




Notice that there are still plural words which can be deceptive

```{r}
ggplot(qs_props_section_type, aes(x = Epigraph, y = Story, 
                      )) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "darkslategray4", high = "gray75") +
  theme(legend.position="none") +
  labs(x = "Epigraph", y = "Story", title = "Word Distributions")
```

```{r}
quakersaints_long |> group_by(younger) |> count(word, sort = TRUE) |>
  mutate(proportion = n / sum(n)) |> select(-n) |>
   pivot_wider(values_from = proportion, names_from = younger) -> qs_props_younger


ggplot(qs_props_younger, aes(x = `TRUE`, y = `FALSE`, 
                      )) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "darkslategray4", high = "gray75") +
  theme(legend.position="none") +
  labs(x = "Younger", y = "Older", title = "Word Distributions")
```




```{r eval=FALSE}
quakersaints_long |> group_by(title) |> count(word, sort = TRUE) |>
  slice_head(n = 10) |> 
   ungroup() |>
  #mutate(word = reorder(word, n)) |>
  
  ggplot(aes( n, word)) +
  geom_col() +
  facet_wrap_paginate(~title, scales = "free_y", ncol =3) +
  theme(strip.text = element_text(
    size = 12, color = "dark green"))
  
```
This is really smushed so I want to split it up and use a smaller 
font.
```{r }
quakersaints_long |> group_by(title) |> count(word, sort = TRUE) |>
  slice_head(n = 10) |>
 mutate(word = reorder(word, n)) |>
  ungroup() -> preprocessed




```

```{r}
  
  preprocessed |>
  filter(title %in% table_of_contents$text[c(4, 6, 8, 10, 12 , 14)]) |> 
   ggplot(aes( n, reorder_within(word,n,  title))) +
  geom_col() +
  facet_wrap(~title, scales = "free_y", ncol =2) +
  scale_y_reordered() +
  theme(strip.text = element_text(
    size = 9, color = "dark green"))


  
```


We could do this multiple times by copy and pasting and modifying 
the list of numbers. 

This is not the best option because all that copy and pasting 
introduces the potential for errors. Also if we make changes, we will
do it in one place.

One solution, make a function.  In this case it will be a very
specific function, not like one for use in other situations.

```{r }
# Chapters is a vector of the chapters to be graphed.
make_faceted_title <- function(chapters ){
     preprocessed |>
      filter(title %in% table_of_contents$text[chapters]) |> 
       ggplot(aes( n, reorder_within(word,n,  title))) +
      geom_col() +
      facet_wrap(~title, scales = "free_y", ncol =2) +
      scale_y_reordered() +
      theme(strip.text = element_text(
        size = 9, color = "dark green"))
}
  
```
```{r}
make_faceted_title(chapters = c(16, 18, 20, 22, 24, 26))
make_faceted_title(seq(from = 28, to = 38, by = 2))
make_faceted_title(seq(from = 40, to = 52, by = 2))
make_faceted_title(seq(from = 54, to = 64, by = 2))
make_faceted_title(seq(from = 66, to = 68, by = 2))
```
We could also make the function even more general by doing things
like passing in the grouping variable and the name of the preprocessed
data set.  We could even extract the grouping variable from the
preprocessed data set.

```{r}
preprocessed |> group_by(title) |>
group_vars()
```

In class/at home: Clean up the document (get rid of unneeded code, make it flow
more smoothly), consider creating a function to reduce repetitive code, and
edit the text to reflect what you did. 

Also do the same kind of data wrangling and more focused analysis on your data. 

I am going to save my trimmed data, simply to maintain consistency at this
point. 

```{r}
save(quakersaints_trimmed, file = here("data/quakersaints_trimmed.rda"))
save(quakersaints_long, file = here("data/quakersaints_long.rda"))

```


```{r}

```

