---
title: "Data reduction of Sentiments"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---


If the file doesn't exist go back and run the qs_nrc file.
```{r message=FALSE}
library(here)
library(dplyr)
library(tidyr)
library(cluster)
library(ggplot2)
library(factoextra)
library(knitr)
```

Let's make a shorter version of the titles.
I realized I needed to do this after making the graphs.

```{r}
if (!exists("qs_nrc" )){
  load(here("data/qs_nrc.rda"))
}
if (!exists("qs_tokens" )){
  load(here("data/qs_tokens.rda"))
}

qs_nrc$title_short <- sub("\\..*","", qs_nrc$section_title)

title_translate <- qs_nrc |> 
  group_by(section_title, title_short) |> 
  summarize(n_tokens = n())

```

Remember from the nrc sentiment analysis "the most common
emotion is trust, followed by anticipation, joy and fear"

```{r}


prop.table(table(qs_nrc$title_short, qs_nrc$sentiment), 2) |>
  round(2) |> kable(caption = "Sentiment by title")
table(qs_nrc$title_short, qs_nrc$sentiment) |>
  prop.table() |>
  as.data.frame() |> pivot_wider(values_from = Freq, names_from = Var2) -> nrcdata



```
I'm going to subset the data so it is just the 
stories and to drop the "negative" and "positive" 
sentiments.

```{r}
nrcdata |> filter(!Var1 %in% c("A TALK ABOUT SAINTS", 
                               "COME-TO-GOOD",
                               "HISTORICAL NOTES",
                               "PREFACE") ) |>
            select(-positive, -negative) ->
  nrcdata
```

I am interested in the question of whether there are patterns
of similarity between the stories.

First I am going to calculate "how far apart" the stories 
are using a distance measure (see separate file explaining this).
For this I use the function `dist()`.  
By reading ?dist you can see that there are 
difference methods to calculate distance:  

"euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski"

The default is euclidean so I'll use that first.

I don't want to use the first column for the calculation
so to keep the labels I am going to creat row.names.

```{r}
nrcdata <- as.data.frame(nrcdata)
rownames(nrcdata) <- nrcdata$Var1
nrcdata <- nrcdata[-1]
```

The distance object shows how far apart the stories
are from each other. This is a distance -- think of it
like inches.  Higher distances means the chapters
are farther apart. A distance of 0 means that they 
are in the same location. 

```{r}
qs_dist <- dist(nrcdata)
#qs_dist 

max(qs_dist)
min(qs_dist)
```
The furthest apart are XXVI and VI.  
XXVI. RICHARD SELLAR AND THE 'MERCIFUL MAN'  
VI. THE SHEPHERD OF PENDLE HILL

The closes are XXIV and XIII  
XIII. MAGNANIMITY
XXIV. THE VICTORY OF AMOR STODDART

When I write up the project The Shepherd of Pendle Hill is 
one I will definitely talk about because it has come up 
in a number of the analyses. 



```{r}

fviz_nbclust(nrcdata, pam, method = "silhouette")+
  theme_classic()
```

So there are two clusters.  This is interesting because
in some of the other analyses it looked like there 
are stories about George Fox and everything else.

```{r}
pam(qs_dist, k = 2, metric = "euclidean", stand = FALSE) -> k2
dd <- cbind(nrcdata, cluster = k2$cluster) 
dd$title_short <- rownames(dd)
dd <- dd |>  left_join(title_translate, by = "title_short")
qs_tokens |>  
  filter(text_type == "Story" & title_type == "main") |>
  group_by(section_title) |>
  summarize(`George Fox` = sum(word == "george_fox"), 
            `Quaker Words` = sum(quaker_word == "Yes"),
            `Younger Readers` = unique(younger)) ->
  quaker_counts
```
```{r}
k2$data <- nrcdata
k2$diss <- qs_dist
fviz_cluster(k2,
    ##palette = c("#00AFBB", "#FC4E07"), # color palette
    ellipse.type = "t", # Concentration ellipse
    repel = TRUE, # Avoid label overplotting (slow)
    ggtheme = theme_classic()
)
```

Story 1 is " 'STIFF AS A TREE, PURE AS A BELL'"
Story 9 is "Strike again"

Stand somewhat outside the ellipses. Both are retellings
(sometimes direct quotations from) of passages in 
the Journal of George Fox.

```{r}
dd |> select(cluster, section_title) |> 
  arrange(cluster, section_title) |>
  kable(caption = "Cluster assignments for stories")
```

```{r}
dd <- dd |> left_join(quaker_counts)
table(dd$cluster, dd$`George Fox`)
table(dd$cluster, dd$`Quaker Words`)

dd |> group_by(cluster) |> 
  summarize(`N of Stories` = n(),
            `Quaker Words` = sum(`Quaker Words` > 0),
            `George Fox` = sum(`George Fox` > 0),
            `For Younger` = sum(`Younger Readers`)) |>
  kable(caption = "Characteristics of Stories")
```

The Stories in Cluster 1 are less likely to include "george_fox" or use Quaker words, but even more dramatic
is that they are less than half as likely to be 
in the appropriate for younger children category.

