---
title: "Curating the data, part 1"
output: html_document
---
```{r}
library(dplyr)
library(stringr)
library(here)
```


This document contains the process of going back to the 
`quakersaints` data and cleaning it up. This way the 
tokenization process will ultimately give us the data that we 
really want to analyze.

It's important to know that this is an _iterative_ process. That
is that you will find yourself going back to the beginning and
making changes or realizing that you left out an important step
that you need later.  So some of the work you will see here
is the product of me having gone through the data set many times.
Just like any kind of creative work, you may need to make 
corrections, reorganize or sometimes go back to the beginning
and make adjustments.

I have also looked at the book itself to understand its structure.

```{r}
# If the file of saved data is not present
if (!file.exists(here("data/quakersaints.rda"))) {
     # download the data from Project Gutenberg
     quakersaints <-  gutenberg_download(c(19605))
     # and save it to that file name.
     save(quakersaints, file="data/quakersaints.rda")
# Otherwise, 
} else {
     # load the data set from the file.
     load(here("data/quakersaints.rda"))
}
```


Looking specifically at this book, I notice first that there are the 
standard Project Gutenberg texts at the beginning. There is also
the title page and a list of other books and dedcation.  These end at
line 81. 
This section is followed by an 
preface which starts on line 82 and  ends on line 182.

Then from 192 to 260 there is a table of contents with blank lines. 
Also * in the title indicates "for younger children." 

Then come the stories. 

I would like to separate each of the stories so that later I can compare
them.  However I really don't want the Table of Contents itself in 
my final data. 

Finally, from line 15189 to the end is also Project Gutenberg content.

Using this information, I can modify the data to get what I want.

A useful first step is to add a variable to the data that identifies
the original row that the data came from.  This will turn out to 
be really helpful for finding information in the original data and
it also is data about the order in which things appear. Later, when
we go back to `unnest_tokens()` this information will be associated
with each token.

Note that because this isn't saving this copy of `quakersaints` 
so if we make a mistake, we can always go back and reload
the original saved file.

```{r}
quakersaints |> mutate(row_number = row_number()) -> quakersaints
```
I can subset the data so that it only includes lines 82 to 15188.
I'm going to call the new data set `quakersaints_trimmed`.

```{r}
quakersaints |> 
  filter(row_number > 81 & row_number < 15189) ->
  quakersaints_trimmed
```

I want to break my text into chapters.  I also can see that there 
are different types of text. For example, the Preface and Historical
Notes are not stories. Also "COME-TO-GOOD" and " A TALK ABOUT SAINTS"
Do not have roman numerals.  A closer look at the text reveals
that these are not stories but are thoughts from the author directly
addressed to the reader.

One thing that I also note about the book is that each story
has a section before the story it self that consists of
powms, quotations and other items. I would like to be able
to separate those out, while keeping track of which epigraph
goes with each story.

There are multiple ways that you can approach this. In ths 
esample I am going to start by working with the table of contents.
Since the titles appear both in the table of contents
and at the beginning of each chapter and epigraph.  I know this because
I looked at the layout of the book in both the data file 
and in the html version of the book on the Project Gutenberg website.
I also note that the title is repeated before both the epigraph and
the man text.

```{r}
table_of_contents <- quakersaints |>
                     filter(row_number >= 190 & row_number <=260) |>
                     filter(text != "")
table_of_contents$text
```

Looking at these results I see that there is work I have to do 
in order to make it possible to match the titles here with the titles
in the rest of the text.

First, I want to take off the "*" characters, but I want to 
create a new variable called `younger` that indicates which
stories were classified that way,
I want to the remove the page numbers from each line, including
the "_page_ vii" from the lines.  

I also notice that there is "white space", probably " " characters,
that I want to remove.
I also want to indicate which of the chapters are not stories 
and which are the sections beore the chapters (Front Matter) and
after them (Back Matter).

To make things easier to read, I'll break out the steps, although
they could be put together using one big `mutate()`. While you are working,
it often makes it easier to check for errors to do this. 

```{r}

# add the indicator for younger audiences
table_of_contents <- table_of_contents |>
                     mutate(younger = substr(text, 1, 1) == "*")
```
Next, let's create a variable that indicates these are titles.
This will be useful later, when we want to separate titles
from text.
```{r}

table_of_contents <- table_of_contents |> mutate(title = "Title")
```



```{r}
# Remove numbers (could also use "[0-9]+")
table_of_contents <- table_of_contents |> 
  mutate(text =gsub("[[:digit:]]+", "", text))
# Deal with a special case
table_of_contents <- table_of_contents |>
  mutate(text = sub("_page_ vii", "", text, fixed = TRUE))
# Remove *
table_of_contents <- table_of_contents |> 
  mutate(text = sub("*", "", text, fixed=TRUE))
# Remove extra white space
table_of_contents <- table_of_contents |> 
  mutate(text =trimws(text))
table_of_contents$text
```
Next, I'll create a variable that indicates whether the title
is for a story, front matter or back matter.
This uses a function called `case_when()` that lets us set up
more complex rules.

```{r}
  table_of_contents <- table_of_contents |>
                mutate(text_type  = case_when (
                  text %in%   c("PREFACE", "A TALK ABOUT SAINTS") ~ 
                                          "Front Matter",
                            text %in% 
                               c("HISTORICAL NOTES", "COME-TO-GOOD") ~
                                    "Back Matter",
                         TRUE ~ "Story")
                  )
```
```{r}
head(table_of_contents)
```



This shows how to do takes some first steps toward combining the above
steps into a longer piped sequence.
```{r eval=FALSE}
table_of_contents <- table_of_contents |>
                     mutate(title = "Title",
                            younger = substr(text, 1, 1) == "*",
                            title = "Title",
                            text_type  = case_when (
                                         text %in%  c("PREFACE", "A TALK ABOUT SAINTS") ~
                                          "Front Matter",
                                          text %in%  c("HISTORICAL NOTES", "COME-TO-GOOD") ~
                                    "Back Matter",
                                          TRUE ~ "Story"),
                            text =gsub("[[:digit:]]+", "", text),
                            text = sub("_page_ vii", "", 
                                       text, fixed = TRUE),
                            text = trimws(
                              sub("*", "", text, fixed=TRUE)
                              ),

                  
                            )
```


Now we can use `table_of_contents` to find the lines in
`quaker_saints`_trimmed that contain titles by matching the
`text` values.  There are (as usual) lots of ways to do this. I'll
do it with a left join.

```{r eval=FALSE}
table_of_contents |>
  left_join(quakersaints_trimmed, by = "text")
```

This gives us a data set with some unexpected results.

First, instead of row_number we have two variables: row_number.x and
row_number.y. These numbers are different because they represent the
row numbers from the table of contents and the row numbers from the 
original text. 
We also have gutenberg_id.x and gutenberg_id.y which is the same,
because both the table of contents and the text came from the same 
original file.  

Also, except for the Preface there are two copies of each title.
This is because the epigraphs and the stories both have the 
titles where they start.

This time we will remove the gutenberg_id and row_number before  doing the 
join.

`title_row` represents the row number where a match to the title in
the table of contents was found.

```{r eval=FALSE}
table_of_contents |>
  select(-gutenberg_id)  |>
  left_join(quakersaints_trimmed, by = "text") |>
  rename(title_row = row_number) ->
  table_of_contents
```

One thing that we can notice is that two of the titles don't
seem to match anything in the `quakersaints_trimmed`:   
XXI. PALE WINDFLOWERS   
XXVII. TWO ROBBER STORIES--WEST AND EAST    

If we look back at the original text (https://www.gutenberg.org/cache/epub/19605/ we can see that in the body
of the book, the titles are slightly different. 

XXI. PALE WIND FLOWERS: OR THE LITTLE PRISON MAID
XXVII. TWO ROBBER STORIES. WEST AND EAST

Probably the simplest thing to do is to chang the table of 
contents version to match that in the text. 

For this it is probably simplest to use base R. 
It took some work to find the exact right punctuation for each.
This changes the specific rows to match exactly. 

```{r}
table_of_contents$text[table_of_contents$text == 
                         "XXI. PALE WINDFLOWERS"] <-
  "XXI. PALE WIND FLOWERS: OR THE LITTLE PRISON MAID"

table_of_contents$text[table_of_contents$text %in% c( 
                         "XXVII. TWO ROBBER STORIES--WEST AND EAST",
                         "XXVII. TWO ROBBER STORIES-WEST AND EAST"
                         )] <-
  "XXVII. TWO ROBBER STORIES. WEST AND EAST"

```

Now I can try the matching again.

```{r }
table_of_contents |>
  rename(toc_row = row_number) |>
  select(-gutenberg_id)  |>
  left_join(quakersaints_trimmed, by = "text") |>
  rename(title_row = row_number) ->
  table_of_contents
```




This still leaves us with two copies of most of  the titles.
We can number these and in most cases use the numbers to indicate whether they
represent epigraphs or the main texts. We will use some
custom code to handle the fact that the Preface does not
have an epigraph.

```{r}
table_of_contents |> 
  group_by(text) |> 
  mutate(title_number = row_number(),
         title_type = case_when(
                         text == "PREFACE" ~ "main",
                         title_number == 1 ~ "epigraph",
                         title_number == 2 ~ "main"
         )) |>
    ungroup() -> 
  table_of_contents
```


Finally, and most important, we want to connect the `table_of_contents`
information to the `quakesaints_trimmed` data.


Again this required some thinking and experimentation to
develop a procedure for. This is because if you just join on "text"
you end up with 4 copies of each title instead of two. 

First I am going create a new variable by copying the value of text
in the  `table_of_contents` data. 

We do not need the row_number from the table of contents in order
to work with the text data.

```{r}
table_of_contents |> 
  mutate(section_title = text) ->
  table_of_contents
```

Now we can join, but instead of using a simple `by` option, we use
the function `join_by` variables that have different names (and more
than one variable at the same time.)
The `arrange()` function is used to ensure that the resulting
data set is in the original order.

```{r}

quakersaints_trimmed |>
  left_join(table_of_contents,
             by = join_by( "text" == "text",
                           "row_number" == "title_row",
                           "gutenberg_id" == "gutenberg_id")) |>
  arrange(row_number) ->
  quakersaints_trimmed
```

Now we have the same 15,107 rows that we started with in 
`quakersaints_trimmed`.  
However, a look at the data shows us that there are many missing
values. Every row that was not a title is missing on most variables.
For one, `title`, we would like to label all of the rest of the 
rows as `text`.  
For most of the others ("younger", "text_type", and "title_type" )
we want every row in the connected text to have the same values.


```{r}
quakersaints_trimmed |> mutate(title = ifelse(is.na(title) == TRUE, 
                                             "Text", "Title")) |>
                        fill(text_type) |> 
                        fill(title_type) |>
                        fill(section_title) |>
                        fill(younger) -> 
  quakersaints_trimmed 
```

I'm going to save `quakersaints_trimmed` so that we can use
the new data set, with all its extra information to tokenize 
and also to compare types of text and specific stories to each other.

```{r savetrimmed}

save(quakersaints_trimmed, file = here("data/quakersaints_trimmed.rda"))

```
