
```{r}
library(ggplot2)
```


# Understanding distance

Normally when people think about distance, they think of 
something like miles or kilometers. 
To measure distance in physical travel we can ask how many
miles is it between Point A and Point B.

But right away there is a question: do we want to know 
how far apart they are in a straight line? Or how many 
miles we would have to drive in a car to get there? 
What about if we walked? Would that be a different number
of miles?  What about if we tried to take the train and/or bus?
How many miles is that? And what would happen if we 
measured distance in time rather than miles? 

In data science we have the same kind of issues.  We can 
want to know how far apart two different observations (rows) are
from each other. This is a surprisingly useful question in 
all kinds of topics and methods.  There are also many different
ways to measure distance. 

Let's look at euclidian distance in R. 

For our first example let's have one variable with two
observations.

```{r}
example1 <- data.frame(var1 = c(5, 10))
dist(example1)
```

As you can see, the distance between the two observations
is 5. To get from 5 to 10 you need to travel 5 units.

What if we have 3 observations? 
```{r}
example2 <- data.frame(var1 = c(5, 10,  7))
dist(example2, diag =TRUE, upper=TRUE)
```
Here we can see that 3 observations means there are 3 
distances:
Observation 1 to Observation 2 is still 5.
Observation 1 to Observation 3 is 2.
Observation 3 to observation 2 is 3.

Make sure you see where those numbers come from. 

Some important things to notice about Euclidean distances are. 

1. The distance between any observation and itself is 0 so 
the display of the distances does not include this information.

2. The distance from observation 1 to observation 2 is the same 
as the distance from observation 2 to observation 1 (in this version of reality). So R only shows each of these one time.  

3. Distances are positive. It's "how many miles apart" 
without information on direction of travel.

Now let's get a little more complicated and have two
variables for each row.

```{r}
example2 <- data.frame(var1 = c(5, 10), var2 = c(1,6))
dist(example2)
```


There is still just one distance between observation 1 and
observation 2. However the Euclidian distance is now given as
(rounded) 7.07.

How did R get this? 

It may help to visualize the data. 

```{r warning=FALSE}
example2 |> ggplot(aes(x = var1, y = var2)) +
  geom_point() +
  geom_line() +
 geom_segment(aes(x = 5, y = 1, xend = 10), color = "darkgray") +
 geom_segment(aes(x = 10, y = 1, yend = 6), color = "darkgray")
```

You could make a right triangle (one angle is 90 degrees) by 
drawing a horizontal line starting from the point at (5, 1) and going to 
(10, 1). That line would be 5 units long (try it and see).

Then make a straight line that goes from the point
at (10, 6) to (10, 1). That would be 6-1 or 5 also. 

We can measure the overall distance by using the Pythagerean Theorem 
which applies to right triangles. The _a_ will be the distance
of the horizontal line and the _b_ will be the distance of the
vertical line. 

$$a^2 + b^2 = c^2$$

So to calculate the distance this we would say 

$$5^2 + 5^2 = distance^2 $$

and then take the square root of the distance squared.

```{r}
(5^2 + 5^2) |> sqrt()
```

And there we can see that it matches what R calculated.

R can also use other methods. For example, New Yorkers
know that you can measure distances in terms of blocks you
need to walk. So from Fifth Avenue and 34th St to 
Park Avenue and 33rd Street, you have to walk 2 blocks,
one down Fifth Avenue to 33rd Street followed by
going on 33rd Street from Fifth Avenue to Park.

```{r}
dist(example2, method = "manhattan")
```

Confirm for yourself that the points are 10 blocks apart.
You can use the graph to help.

There are also other options.

What about bigger data sets, with more than 2 columns and 
many rows? R can handle that too! 

A related concept is _similarity_.  In similarity,
the 
```{r}
library(proxy)
simil(example2)
```
With this simple data there is no similarity between
the rows. 

Let's try something more complicated.

```{r}

example3 <- data.frame(var1 = c(5, 10,  7),
                       var2 = c(1, 6, 7), 
                       var3 = c(5, 10, 10))
dist(example3)
simil(example3, by_rows = TRUE)


```

Here there are three rows and three columns.  
The distance matrix is giving us a measure of the distances
between the three points (they are the x, y and z values so 
it's distance in 3-dimensional space).  Bigger numbers
mean further distances.

With similarity bigger numbers means more similar. So
the most similar rows are the second and third. This 
makes sense because those two rows have a common value (10)
and two of the values (6 and 7 in the second column) are
close to each other.

We can also compute column similarity. 

```{r}
simil(example3, by_rows = FALSE)
```

When using clustering methods you are using specific kinds
of distance and similarity measures. The clustering methods
generally work by putting similar observations into 
clusters using some criteria (which can vary) to decide
what the best solution is.
