---
title: "Poster Supplement"
output: html_document

---

This file contains supplemental information for my poster.
If I want, I can put this on a web page and add a QR code
to my poster leading to this document.

## George Fox and others

George Fox was the founder of Quakerism and
what is presumably his last name is the top word in 5 stories. Of 
course it is possible that some instances may be a reference to the
animal or to some other person named fox. 

What if we look at how often the word fox appears in all of the
stories.
```{r}
# If the file of saved data is not present
if (!file.exists(here("data/quakersaints_trimmed.rda"))) {
      message("quakersaints_trimmed needs to be recreated")
# Otherwise, 
} else {
     # load the data set from the file.
     load(here("data/quakersaints_trimmed.rda"))
}
```

```{r}
quakersaints_trimmed |>
  mutate(text = tolower(text)) |>
  mutate(text = str_replace_all(text,
             c("george fox" = "george_fox",
               "g. fox" = "george_fox", 
               "robert fowler" = "robert_fowler",
               "master robert fowler" = "robert_fowler",
               "james nayler" = "james_nayler",
               "judge fell" = "judge_fell", 
               "john audland" = "john_audland",
               "meeting-house" = "meeting_house",
               "meeting house" = "meeting_house" ,
               "steeple-house" = "steeple_house",
               "steeplehouse" = "steeple_house",
               "valiant sixty" = "valiant_sixty", 
               "edward burrough" = "edward_burrough",
               "e. burrough" = "edward_burrough",
               "john camm" = "john_camm",
               "sir edward" =  "sir_edward",
               "grand vizier"  = "grand_vizier",
               "mary fisher"  = "mary_fisher",
               "mistress fell" = "margaret_fell",
               "margaret fell" = "margaret_fell",
               "margaret fox" = "margaret_fell",
               "william dewsbury" = "william_dewsbury",
               "james parnell"   = "james_parnell",
               "william savery" = "william_savery",
               "leonard fell"  = "leonard_fell",
               "preston patrick" = "preston_patrick",
               "swarthmoor hall" = "swarthmoor_hall",
               "lord power"  = "lord_power",
               "stephen grellet" = "stephen_grellet",
               "etienne de grellet" = "stephen_grellet",
               "tanner savery" = "william_savery",
               "mary savery" = "mary_savery",
               "mrs savery" = "mary_savery",
               "george robinson" = "george_robinson",
              "richard sellar" =  "richard_sellar",
              "shepherd of pendle" = "shepherd_of_pendle",
              "rough moll" = "rough_moll" #,
              #"the man in leather breeches" = "george_fox",
              # "the shepherd of pendle hill" = "george_fox"
                                  )))|>
  # remove an empty rows
  filter(!is.na(text))  -> quakersaints_trimmed2

```
```{r}
quakersaints_trimmed2 |> 
  unnest_tokens(output=word, input = text, token = "words") ->
   qs_tokens

```
```{r}
 qs_tokens |> anti_join(stop_words) -> qs_tokens

```
```{r}
save(qs_tokens, file = here("data/qs_tokens.rda"))
```



```{r}
qs_tokens |> filter(text_type == "Story" & title_type =="main"
                    ) |>
             group_by(section_title) |>
             summarize(uses = sum(word == "george_fox")) |>
             kable(caption = "Presence of 'george_fox' by story")
```


So despite george_fox being the most common word, there are there are 5 stories
which don't use it at all. Among those which have the word, there 
are big differences. The story "Taming the Tiger" has the most uses,
43.

One thing that is striking to me is that "THE MAN IN LEATHER BREECHES" 
and "THE SHEPHERD OF PENDLE HILL" don't mention Fox, because 
both of those descriptions refer to Fox.


## Quaker Words 

Over time, working with this book, I concluded it was important
to track "Quaker words". The solution I came up with was to 
add a variable indicating if a token was one of these words. (I also
considered treating them as stop words.)
This is an example of how any text you are working with may have
very specific issues that may need to be considered.

```{r}
quaker_words <- c("thou", "thee", "thy", "art", "hast", "thine")
qs_tokens |> mutate(quaker_word = ifelse(word %in% quaker_words, 
                                         "Yes", "No")) -> qs_tokens

```

