---
title: "Sentiment Analysis using the Bing Lexicon"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(here)
library(dplyr)
library(tidytext)
library(tidyr)
library(knitr)
library(ggplot2)
```

Sentiment analysis allows us to use standard lexicons to explore the
sentiments expressed in texts.

This is a basic sentiment analysis. Notice that it assumes
that two data sets have been created and saved in the data folder.
I use `knitr::knit_exit()` to stop knitting if either of
the data sets are not present.

```{r}
if (file.exists(here("data/qs_tokens.rda"))) {
     load(here("data/qs_tokens.rda")) 
} else {
  print("qs_tokens not present.")
  knitr::knit_exit()
}
```

We will use the tokenized data (one word per line). In this data you should
also have a row number variable and a variable that defines which
chapter/story/section the word is in. We will be using those variables
to help us to explore sentiments in the text.

In Bing, some words are both positive and negative so it is 
possible for a word to have two matches. Therefore the 
relationship="many-to-many" option is used.

A "join" is a method of putting two data frames together. An
inner join keeps only those rows that are present in both data sets.
The "by" option indicates the variable or variables that will
be used to match the rows in the two data sets.

This data is only going to include words that have sentiment values.

```{r}
bing <- get_sentiments("bing")
bing |> inner_join( qs_tokens, by = "word", multiple = "all",
                    relationship = "many-to-many") -> qs_bing
```

```{r}
table(qs_bing$sentiment)
table(qs_bing$sentiment) |> prop.table() |>round(2)
```

Overall the lannguage is more negative than positive. However, in
the Bing lexicon 70% of the entries are negative, so we wouldn't
say that the words distribution in the Quaker stories is necessarily
overly negative.

Interestingly, when stories appropriate for younger audiences
are compared to those that are not, the younger-oriented 
stories are, if anything, slightly more negative although 
there are only small differences..

## Stories for younger versus older readers

```{r}
table(qs_bing$sentiment, qs_bing$younger) |> prop.table(2) |> round(2)

```

## Sentiments by type of text

```{r}
table(qs_bing$sentiment, qs_bing$text_type) |> prop.table(2) |> round(2)

```

Interestingly the stories are more negative than the other materials.

This does tend to support the idea that the epigraphs are more flowery.

Based on this, I'm going to redo the earlier comparisons of young and
old with just stories.

```{r}
qs_bing |> filter(text_type == "Story") -> qs_stories_bing
table(qs_stories_bing$sentiment, qs_stories_bing$younger) |> 
  prop.table(2) |> round(2)

```

The results are essentially the same.

How much variation is there from story to story?

Some chapters tend to have more negatives words compared to other
stories. For example the 'THE MAN IN LEATHER BREECHES' story has a
**0.76** negative word ratio compared to **0.24** positive word ratio.

```{r}
table( qs_stories_bing$title, qs_stories_bing$sentiment) |> 
  prop.table(1) |> round(2) |>
  kable(caption = "Proportion of words with sentiments")
```

Let's use a different method and sort the results.

```{r}
 qs_stories_bing |> group_by(section_title) |> 
  summarize(Positive = round(sum(sentiment == "positive")/n(), 2),
            Negative = round(sum(sentiment == "negative")/n(), 2)) |>
  arrange(Positive) |> kable(caption = "Sentiment breakdown by story")
```

Visualizing this might be better. This is one possible way.
For yours make sure to add labels. In this case I might also want to add
a line highlighting .5 or to change the tick marks to help visualize
that. Do what makes sense for your data.

```{r fig.height=5}
 qs_stories_bing |> group_by(section_title) |> 
  summarize(Positive = round(sum(sentiment == "positive")/n(), 2)) |>
  arrange(Positive) |> 
  ggplot(aes(x=reorder(section_title, Positive), y =Positive)) +
  geom_point(size = 4) +
   geom_segment(aes(x=section_title, xend=section_title, y=0, yend=Positive)) +
  coord_flip() +
  labs(title = "Stories ranked by proportion of positive words")


```

```{r fig.height=5}
 qs_bing |> group_by(section_title) |> 
  summarize(Negative = round(sum(sentiment == "negative")/n(), 2)) |>
  arrange(Negative) |> 
  ggplot(aes(x=reorder(section_title, Negative), y =Negative)) +
  geom_point(size = 4) +
   geom_segment(aes(x=section_title, xend=section_title, y=0, yend=Negative)) +
  coord_flip() +
  labs(title = "Stories ranked by proportion of negative words")


```


Another thing to look at will be how sentiments change during the
stories.

How do we have to fix this to make it meaningful?
we can change the visualization of the charts by giving a different graph to each story.

```{r fig.height=8}
 qs_stories_bing |> 
  # Select just the first 10 stories
  filter(title %in% unique(title)[1:10]) |>
  # I'm going to make groups of 10 rows to do together.
  # I could use modulo instead.
  mutate(row_10 = round(row_number/10), 0) |>
  arrange(row_10) |>
  group_by(title, row_10) |>
  summarize(negative = round(sum(sentiment == "negative")/n(), 2)) |>
ggplot( aes(x= row_10, y = negative)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~title, ncol = 2, scales = "free_x") +
  labs(title = "Negative sentiments over the course of the stories",
      subtitle = "Stories 1 to 10")
```

```{r fig.height=8}
 qs_stories_bing |> 
  filter(section_title %in% unique(section_title)[11:20]) |>
  arrange(row_number) |>
  group_by(section_title, row_number) |>
  summarize(negative = round(sum(sentiment == "negative")/n(), 2)) |>
ggplot( aes(x= row_number, y = negative)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~section_title, ncol = 2, scales = "free_x")
```



```{r fig.height=8}

 qs_stories_bing |> 
  # Select just the first 10 stories
  filter(section_title %in% unique(section_title)[11:20]) |>
  # I'm going to make groups of 10 rows to do together.
  # I could use modulo instead.
  mutate(row_10 = round(row_number/10), 0) |>
  arrange(row_10) |>
  group_by(section_title, row_10) |>
  summarize(negative = round(sum(sentiment == "negative")/n(), 2)) |>
ggplot( aes(x= row_10, y = negative)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~section_title, ncol = 2, scales = "free_x") +
  labs(title = "Negative sentiments over the course of the stories",
      subtitle = "Stories 11 to 20")
```

```{r fig.height=9}

 qs_stories_bing |> 
  # Select just the first 10 stories
  filter(section_title %in% unique(section_title)[21:32]) |>
  # I'm going to make groups of 10 rows to do together.
  # I could use modulo instead.
  mutate(row_10 = round(row_number/10), 0) |>
  arrange(row_10) |>
  group_by(section_title, row_10) |>
  summarize(negative = round(sum(sentiment == "negative")/n(), 2)
            ) |>
ggplot( aes(x= row_10, y = negative)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~section_title, ncol = 2, scales = "free_x") +
  labs(title = "Negative sentiments over the course of the stories",
      subtitle = "Stories 21 to 32")
```

There are many additional ways we could analyze these data.
What is useful is going to depend on your text and the questions
you are interested in.