---
title: "Sentiment Analysis using the nrc Lexicon"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(here)
library(dplyr)
library(tidytext)
library(tidyr)
library(knitr)
library(ggplot2)
```

Sentiment analysis allows us to use standard lexicons to explore the
sentiments expressed in texts.

This is a basic sentiment analysis. Notice that it assumes
that two data sets have been created and saved in the data folder.
I use `knitr::knit_exit()` to stop knitting if either of
the data sets are not present.

```{r}
if (file.exists(here("data/qs_tokens.rda"))) {
     load(here("data/qs_tokens.rda")) 
} else {
  print("qs_tokens not present.")
  knitr::knit_exit()
}
```

We will use the tokenized data (one word per line). In this data you should
also have a row number variable and a variable that defines which
chapter/story/section the word is in. We will be using those variables
to help us to explore sentiments in the text.


This website tells you more about the nrc lexicon.
https://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm

This explains more about how it was created.
How Mechanical Turkers Crowdsourced a Huge Lexicon of Links Between Words and Emotion. n.d. Retrieved February 17, 2026. https://www.technologyreview.com/2013/09/05/176552/how-mechanical-turkers-crowdsourced-a-huge-lexicon-of-links-between-words-and-emotion/.


Mohammad, Saif. 2011. “From Once Upon a Time to Happily Ever After: Tracking Emotions in Novels and Fairy Tales.” Pp. 105–14 in Proceedings of the 5th ACL-HLT Workshop on Language Technology for Cultural Heritage, Social Sciences, and Humanities, edited by K. Zervanou and P. Lendvai. Portland, OR, USA: Association for Computational Linguistics.


NRC distinquishes "sentiment" (positive and negative) from
emotion (specific feelings). For emotions, descriptive terms such as
"trust", "fear", "sadness" and "anger" are assigned to words.
In some cases a word may be assigned multiple emotions.

A "join" is a method of putting two data frames together. An
inner join keeps only those rows that are present in both data sets.
The "by" option indicates the variable or variables that will
be used to match the rows in the two data sets.

Because the same emotions and sentiments are assigned to many words
AND specific words are assigned multiple emotions we use what is called
a many-to-many option when carrying out the join.  

This data is only going to include words that have sentiment values.

```{r}
nrc <- get_sentiments("nrc")
nrc |> inner_join(qs_tokens, by = "word", multiple = "all",
                   relationship = "many-to-many") -> qs_nrc
```


```{r}

```

Overall the words in our files are more positive than negative. 

Of course there are many questions we can ask about these distributions.

First, descriptively, how many are present? 

```{r}
qs_nrc |> 
  group_by(sentiment) |> 
  summarize(number = n())
```
In this case we have more positive than negative
sentiments and the most common emotion is trust, followed by
anticipation, joy and fear. 

```{r}
qs_nrc |> 
  group_by(sentiment) |> 
  summarize(number = n(), 
            n_younger = round(sum(younger == TRUE)/number, 2), 
            n_older = round(sum(younger== FALSE)/number, 2)
           )
```

If we look at the proportion of sentiment and emotion words
that are associated with words in older and younger stories
we see that consistnently the majority are in stories for 
younger readers.  

However, we need to be cautious about interpreting this too much
because there are more total words in the younger stories. 

Therefore we would expect that there are more sentiment-emotion 
words in the stories for younger readers, simply by chance. 
This is something that could be explored in a more complex way.

```{r}
table(qs_tokens$younger)
table(qs_tokens$younger) |> prop.table() |> round(2)
```

We could look at what the most common emotion in each story is.
In this case we could just focus on the emotion words.

```{r}
qs_nrc |> filter (!sentiment %in% c("positive", "negative") ) |>
  group_by(section_title, sentiment) |> 
  summarize(n = n()) |>
  ungroup() |>
  group_by(section_title) |>
    arrange(n) |>
  slice_tail(n =1) |>
  kable(caption = "Most common emotion in each story")
```

It may not be surprising that trust is the most common emotion in
many of the stories. In the lexicon trust is matched with
494 different words in our data set.  Twenty of them are shown.

```{r}
qs_nrc |> filter(sentiment == "trust") -> trustwords
length(unique(trustwords$word))

head(unique(trustwords$word), 20)

```

```{r}
save(qs_nrc, file = here("data/qs_nrc.rda"))
```

