Stemming is the process of reducing a word to it's "stem" which then allows us to do things such as treat "walk", "walked" and "walking" as the same words.
This gives us better or at least different understanding of the most frequent words.

Porter2 Snowball stemming  

Review this page 
[https://snowballstem.org/algorithms/english/stemmer.html]

```{r}
library(SnowballC)

```

```{r}
wordStem(c("win", "winning", "winner"))
wordStem(c("call", "calling", "called"))
```
```{r}
getStemLanguages()
```

This will generate a data frame of words and stems for specific languages.

```{r}

lang <- "english"
load(system.file("words", paste0(lang, ".RData"), package="SnowballC"))

```
Do you think they are resonable? 


Also load your long data.

Use the wordStem(words, language = "english") function to create a new version of your long data with the stemmed words.   
(Note that this function is in the quatenda package with the name tokens_wordstem or dfm_wordstem for dfm objects).


```{r}
wordStem(quakersaints_long$word) -> quakersaints_long$word_stemmed

# quakersaints_long |> mutate(word_stemmed = wordStem(word)) -> quakersaints_long
```

Now rerun you analysis of the top words.  Do any rankings change when you 
use the stemmed words instead of the unstemmed words? 
