Stemming is the process of reducing a word to it's "stem" which then allows us to do things such as treat "walk", "walked" and "walking" as the same words.
This gives us better or at least different understanding of the most frequent words.

Porter2 Snowball stemming  

Review this page 
[https://snowballstem.org/algorithms/english/stemmer.html]

```{r}
library(SnowballC)

```

```{r}
wordStem(c("win", "winning", "winner"))
wordStem(c("call", "calling", "called"))
```
```{r}
getStemLanguages()
```

This will generate a data frame of words and stems for specific languages.

```{r}

lang <- "english"
load(system.file("words", paste0(lang, ".RData"), package="SnowballC"))

```
Do you think they are resonable? 


Also load your long data.

Use the wordStem(words, language = "english") function to create a new version of your long data with the stemmed words.   
(Note that this function is in the quatenda package with the name tokens_wordstem or dfm_wordstem for dfm objects).


```{r}
qs_tokens |> mutate(word_stemmed = wordStem(word)) -> qs_tokens
save(qs_tokens, file = here("data/qs_tokens.rda"))

# quakersaints_long |> mutate(word_stemmed = wordStem(word)) -> quakersaints_long
```

Now rerun you analysis of the top words.  Do any rankings change when you 
use the stemmed words instead of the unstemmed words? 


```{r}
qs_tokens |> count(word_stemmed, sort = TRUE) |>
  filter(n > 200) |>
  mutate(word_stemmed = reorder(word_stemmed, n)) |>
  ggplot(aes(n, word_stemmed)) +
  geom_col() +
  labs(y = NULL)
```

```{r}
qs_tokens |> count(word, sort = TRUE) |>
  filter(n > 200) |>
  mutate(word = reorder(word, n)) |>
     
  ggplot(aes(x = n, y = word)) +
  geom_col() +
  labs(y = NULL)
```

```{r}
qs_tokens |> count(word_stemmed, sort = TRUE) |>
  filter(n > 200) |>
  mutate(word_stemmed = reorder(word_stemmed, n)) 
```

Stemming the words increases the
ranking of some of the words and
decreases others.
For example, unstemmed, fox and god were the top two words, but 
with the stemmed data fox moves to
the second position and god moves to 
fifth position.  The string
"dai" is the most common word in
the stemmed data, while "day" 
was in the fourth position in
the unstemmed data.

Time is also common in both. Maybe
there is something going on with
ideas about time (time and day are both about time). We should look
more closely at this.

