---
title: "Tokenization part 2"
output: html_document
---

```{r}
  library(dplyr)
  library(here)
  library(tidytext)
  library(knitr)
  library(tidyr)
  library(ggplot2)
  library(stringr)
```
This file works on the trimmed and cleaned data as created n
curation_toc.

```{r}
# If the file of saved data is not present
if (!file.exists(here("data/quakersaints_trimmed.rda"))) {
      message("quakersaints_trimmed needs to be recreated")
# Otherwise, 
} else {
     # load the data set from the file.
     load(here("data/quakersaints_trimmed.rda"))
}
```





```{r}
quakersaints_trimmed |> 
  unnest_tokens(output=word, input = text, token = "words") ->
   qs_tokens

```


Once again we will remove the standard stop words.

```{r}
 qs_tokens |> anti_join(stop_words) -> qs_tokens

```

```{r}
qs_tokens |> count(word, sort = TRUE)
```

Unique words

```{r}
unique(qs_tokens$word) |> length()
```




## Some analysis

Now we can look at some interesting questions. For example, 
what is the most common word in each story main text.

```{r}
qs_tokens |> filter(text_type == "Story" & title_type =="main") |>
             group_by(section_title, word) |>
             summarize(uses = n()) |>
             group_by(section_title) |>
             arrange(uses) |>
             slice_tail(n = 1) |>
             kable(caption = "Most common word in each story")
```

Even a list this simple is interesting. For example, most of the 
top words are names. Since this is a book titled "Quaker Saints" it is
not surprising that
the main focus of each chapter is on an individual.


George Fox was the founder of Quakerism and
what is presumably his last name is the top word in 5 stories. Of 
course it is possible that some instances may be a reference to the
animal or to some other person named fox. 

What if we look at how often the word fox appears in all of the
stories.

```{r}
qs_tokens |> filter(text_type == "Story" & title_type =="main"
                    ) |>
             group_by(section_title) |>
             summarize(uses = sum(word == "george_fox")) |>
             kable(caption = "Presence of 'fox' by story")
```


So despite fox being the most common word, there are there are 5 stories
which don't use it at all. Among those which have the word, there 
are big differences. The story "Taming the Tiger" has the most uses,
43. Later, I might want to look more closely at whether there is 
anything correlated with how often the word is used.

One thing that is striking to me is that "THE MAN IN LEATHER BREECHES" 
and "THE SHEPHERD OF PENDLE HILL" don't mention Fox, because 
both of those descriptions refer to Fox.

_What about the Epigraphs?_

```{r}
qs_tokens |> filter(title_type == "epigraph" & text_type == "Story") |>
             group_by(section_title, word) |>
             summarize(uses = n()) |>
             ungroup() |>
             group_by(section_title) |>
             arrange(uses) |>
             slice_tail(n = 1) |>
             kable(caption = "Most common word in each epigraph")
```

Here we see that the words are largely different. With the 
exception of "fox" they are not names. There are also some unusual
(in modern times) words such at "thy" and "thee".  These words
are "second person singular" equivalent to "yours" and "you",
which in modern English are used for both one person and groups 
of people, and both formally and informally.  You may 
know other languages (such as Spanish and French) where the distinction
between formal and informal is used.  Quakers maintained
this distinction, and used the informal for everyone to reflect
their belief in the spiritial equality of all.  For some, this
was maintained well into the twentieth century. 

There are many more spiritual or religious terms such as god, spirt and
spiritual.

Fox's writings are the source of a number of the quotations in the
epigraphs; the references to him are _by_ him, not _about_ him.

This is a good start but as you can see if you look at the head of
qs_tokens there are still a number of smaller issues. For example:

1. there are some words that either begin with "_" or end with "_". (If you
sort the qs_tokens data by word you can see these.)  
2.There
are also words such as "charles's"  wheree mightwant to remove the
"'s".   

3. There are also a lot of numbers; some look like years, but others
are harder to interpret.   

Before moving to further analysis I am going to deal with the first
two issues.  And because these changes may reveal some new stop
words I need to recheck those.

```{r}
qs_tokens |> mutate(word = case_when(
                   str_sub(word, 1, 1) == "_"  ~ gsub("_", "" , word),
              str_sub(word, -1, -1) == "_"  ~ gsub("_", "" , word),
              str_sub(word, -2, -1) == "'s"  ~ gsub("'s", "" , word),
              TRUE ~ word)
)  |> anti_join(stop_words) ->  qs_tokens

```

Over time, working with this book, I concluded it was important
to track "Quaker words". The solution I came up with was to 
add a variable indicating if a token was one of these words. (I also
considered treating them as stop words.)
This is an example of how any text you are working with may have
very specific issues that may need to be considered.

```{r}
quaker_words <- c("thou", "thee", "thy", "art", "hast", "thine", "wilt")
qs_tokens |> mutate(quaker_word = ifelse(word %in% quaker_words, 
                                         "Yes", "No")) -> qs_tokens

```

```{r}
qs_tokens |> filter(text_type == "Story" & title_type =="main"
                    ) |>
             group_by(section_title) |>
             summarize(uses = sum(quaker_word == "Yes")) |>
             kable(caption = "Presence of 'fox' by story")
```
Let's save the tokenized data with the new variable.

```{r}
save(qs_tokens, file = here("data/qs_tokens.rda"))
```

```{r}

```



So even though we have basic data now, we still may
need to do additional curation because other issues may 
appear as we do the analysis.
