---
title: "Topic Models"
output: html_document
---
```{r}
library(here)
library(dplyr)
library(tidytext)
library(tidyr)
library(knitr)
library(ggplot2)
library(tm)
library(topicmodels)
```

To do this analysis in this file, you will need to copy your saved 
tokenized data. 

In this project we are going to take a first look at a method of 
analysis called *topic models*. 
Topic models are a kind of machine learning that analyzes 

# Document-Term Matrix

So far we have been focused on *words* as our units.  For example, in the 
sentiment analysis each row represented one word, and there was a column that
represented the specific document (story, chapter, book, etc).   

In this analysis we will change perspective: the document will represent the
observation, so each document will will be a row.

The columns will represent "terms."  In this case the terms will be individual
words.   So each word acts like a variable.  The values in the column will 
be the frequency of that term in the specific document.  

These usually end up very wide (because there are a lot of words) and
relatively short (because there are not that many documents compared
to the number of words).

```{r}
fake_data_dtm <- data.frame(
  "Document" = c("Book 1", "Book 2", "Book 3"),
  "Cat" = c(0, 4, 1),
  "Sky" = c(0, 0, 1),
  "House" = c(3, 1, 1)
)
fake_data_dtm


```

Notice that there are three 0s.  This is important 
later! This happens when there are words that 
do not appear in a given document.

```{r}
fake_data_long <- data.frame(
  "Word" = c("Cat", "Cat", "Sky", "House", "House", "House"),
  "n" = c(4, 1, 1, 3, 1, 1),
  "Document" = c("Book 2", "Book 3", "Book 3", "Book 1", "Book 2", "Book 3")
)
fake_data_long
```

```{r}
cast_dtm(fake_data_long, Document, Word,  n) -> fake_dtm

fake_dtm$v
fake_dtm
dimnames(fake_dtm)
```
Notice that we get two dimenisons, one for the words and 
one for the terms. 

Looking above you can remind yourself that we had 
three 0s and 6 other numbers.   

Term-data matrices are often *sparse* because of these 
0s.  You can see the two dfferent sparsity measures.

In terms of computing, we often don't want to have to take
up space for all the 0s so we change (cast) the 
matrix ti a "sparse" matrix.

```{r}
cast_sparse(fake_data_long,Document,  Word,  n) -> fake_sparse
fake_sparse

```

There are a few different packages that do this analysis. We are going to 
using the `tm` packages.

```{r}
load(here("data/qs_tokens.rda"))

qs_tokens |> 
  filter(text_type == "Story", 
         title == "Text",
         title_type == "main") ->
  qs_tokens_stories

qs_tokens_stories |> group_by(section_title, word) |> 
  summarize(n = n()) |>
cast_dtm( term = word, document=section_title, n) -> quakersaints_dtm
```

Now I have a "large DocumentTermMatrix" in my environment.

# Latent Dirichlet allocation (LDA)

LDA "treats each document as a mixture of topics, and each topic as a mixture of words" (chapter 6).

This means that each document can have multiple topics.  
Each of the topics is made up of a mixture of words/terms.


Just like with cluster analysis, we need to provide the parameter _k_  which is the number of topics to be 
identified.   In this sense topic models depend on the user, 
so you must have 
some judgement.  Often what we do is to try various values of _k_ to find the
best solution. 

The algorithm finds the best solution for a given number of topics. 

*This means that you can get different answers each 
time you run the analysis!!!*

But a solid analysis should give you pretty simlar 
answers.

It requires a random number to start; we can specify a specific value which
will let us reproduce the same exact anaysis if we want.
However, you should not do this all the time, but 
rather use a "random start." 

```{r}
library(random)
seed <- randomNumbers(n=1, col = 1)
seed
```


```{r}
qs_lda_2 <-
  LDA( quakersaints_dtm, k = 2, control = list(seed= seed))
qs_lda_2 
```
```{r}
qs_lda_3 <-
  LDA( quakersaints_dtm, k = 3, control = list(seed= seed))
qs_lda_3 
```
```{r}
qs_lda_4 <-
  LDA( quakersaints_dtm, k = 4, control = list(seed= seed))
qs_lda_4 
```

## For your data try starting with several different seeds. Do the results change?

Also try different numbers of topics.

The result object `qs_lda` contains a lot of different information. 
We can see that the word assignments are there in there. But how to 
access them?

We can use the `tidy()` function to make them more accessible.

```{r}
qs_topics_2 <- tidy(qs_lda_2, matrix = "beta")
head(qs_topics_2)
```

```{r}
qs_topics_3 <- tidy(qs_lda_3, matrix = "beta")
head(qs_topics_3)
```

```{r}
qs_topics_4 <- tidy(qs_lda_4, matrix = "beta")
head(qs_topics_4)
```

Notice that each row is a topic-term. That is, there is a row for each
combination of topic and word.  Since we set k =2 there are two rows per
word. with a higher k we will see k rows per term.




The `beta` colum represents the probability that the word was generated
for that *topic*.  That means we can find the words with the highest
probability for the topic.  

These are (usually) in scientific notation. 
So a bigger number after the e means more 0s 
between the decimal and the number, that is 
a bigger numbr afterthe e means that the probability
is smaller.

Let's look at the top 10 words for each topic.
That is the 10 words with the highest probabilities.

```{r}
qs_top_terms_2 <- qs_topics_2 |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() %>%
  arrange(topic, -beta)

qs_top_terms_2 |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```

```{r}
qs_top_terms_3 <- qs_topics_3 |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() %>%
  arrange(topic, -beta)

qs_top_terms_3 |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```

Notice that the same words can be in multiple topics (george, god, day).


```{r}
qs_top_terms_4 <- qs_topics_4 |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() %>%
  arrange(topic, -beta)

qs_top_terms_4 |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```


The following approach uses the ratio of the probabilities to find 
the words that _differentiate the topics_.

```{r}
beta_wide_2 <- qs_topics_2 |>
  mutate(topic = paste0("topic", topic)) |>
  pivot_wider(names_from = topic, values_from = beta) |>
  filter(topic1 > .001 | topic2 > .001) |>
  mutate(log_ratio = log2(topic2 / topic1))

beta_wide_2 |> slice_max(order_by = log_ratio, n = 10)
beta_wide_2 |> slice_min(order_by = log_ratio, n = 10)

```


```{r}
beta_wide_3 <- qs_topics_3 |>
  mutate(topic = paste0("topic", topic)) |>
  pivot_wider(names_from = topic, values_from = beta) |>
  filter(topic1 > .001 | topic2 > .001) |>
  mutate(log_ratio = log2(topic2 / topic1))

beta_wide_3 |> slice_max(order_by = log_ratio, n = 10)
beta_wide_3 |> slice_min(order_by = log_ratio, n = 10)

```


```{r}
beta_wide_4 <- qs_topics_4 |>
  mutate(topic = paste0("topic", topic)) |>
  pivot_wider(names_from = topic, values_from = beta) |>
  filter(topic1 > .001 | topic2 > .001) |>
  mutate(log_ratio = log2(topic2 / topic1))

beta_wide_4 |> slice_max(order_by = log_ratio, n = 10)
beta_wide_4 |> slice_min(order_by = log_ratio, n = 10)

```



## Guessing the topic

The above looks at the words associated with a topic.  We can also look
at the _probability we would guess the right topic_ based on the word. This uses
gamma.  

```{r}
qs_documents_2 <- tidy(qs_lda_2, matrix = "gamma")
qs_documents_2
```


```{r}
qs_documents_3 <- tidy(qs_lda_3, matrix = "gamma")
qs_documents_3
```

```{r}
qs_documents_4 <- tidy(qs_lda_4, matrix = "gamma")
qs_documents_4
```

```{r}
top_terms_3 <- qs_topics_3 |>
  group_by(topic) |>
  slice_max(beta, n = 5) |>
  ungroup() |>
  arrange(topic, -beta)

top_terms_3

top_terms_3 %>%
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()

```
We can see that some words differentiate the topics and others do not.


For your analysis try different numbers of topics (these can include large
values of k, don't be afraid to choose 10 or maybe even more. )

Write up a paragraph explaining what you learned about your documents.


```{r}

qs_lda_10 <-
  LDA( quakersaints_dtm, k = 10, control = list(seed = 1234))
qs_topics_10 <- tidy(qs_lda_10, matrix = "beta")
head(qs_topics_10,15)
```

```{r}
qs_top_terms_10 <- qs_topics_10 |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() %>%
  arrange(topic, -beta)

qs_top_terms_10 |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```
Is there a way to figure out the best number 
of topics?  That is, that fit the data the best? 

There is no one clear solution for this. 

One common way

```{r}
library(stm)
ldatuning.metrics <- FindTopicsNumber(quakersaints_dtm, topics = seq(from = 2, to = 15, by = 1), metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"), method = "Gibbs", control = list(seed = 77), mc.cores = 2L, verbose = TRUE
)
```

