---
title: "An alternative approach"
output: html_document
---
```{r}
library(dplyr)
library(stringr)
library(here)
library(tidytext)
library(tidyr)
```

```{r}
if (file.exists(here("data/quakersaints_trimmed.rda"))) {
     load(here("data/quakersaints_trimmed.rda")) 
} else {
  print("quakersaints_trimmed not present.")
  knitr::knit_exit()
}
```
```{r}


quakersaints_trimmed |> 
  unnest_tokens(output=word, input = text, token = "words") ->
   qs_tokens

```

`

Another approach that focuses on highlighting differences, without
using stop words is the calculation of the `tf_idf` for each word
in each document. This stands for Term Frequency-Inverse-Document
Frequency.  This is less complicated than it sounds.

Step 1: What does is ask how many of the stories does the term 
appear in.  What proportion of stories is that? 

```{r}
qs_tokens |> 
  # get the number of times each word appears by item 
  count(section_title, word, sort = TRUE)  -> 
  tf
```
One thing that is interesting here is that because we are using the
words present in a text, we don't actually have the words _not_
present.  Therefore, values_fill = 0 is used to add the 0s.
```{r}
tf |> pivot_wider(id_cols = section_title, names_from = word,
                  values_from = n, values_fill = 0) -> tfm
```

However, let's go back to the `tf` data 
For each word we want to find out how many documents it is used in.

```{r}
total_documents <- unique(tf$section_title) |> length()

tf |>
  group_by(word)  |>
 summarize(n_documents = n(), 
         prop_documents  = n_documents/total_documents,
         ln_prop = log(prop_documents)) ->
  tf_weights

```

We use the log (natural log) here which means that if the proportion
is 1 (that is the word is in all of the stories) we get a tf_weight of 0.

